---
title: "enterotyping"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{enterotyping}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

A workflow for identifying enterotypes based on the relative abundance of gut microbiota was implemented refereed on the reports of Arumugam[^2]

[^1]: Arumugam, M., Raes, J., Pelletier, E., Le Paslier, D., Yamada, T., Mende, D. R., ... & Bork, P. (2011). Enterotypes of the human gut microbiome. nature, 473(7346), 174-180.

```{r setup}
library(mbOmic)
library(data.table)
```

First of all, the dataset of microbiota relative abundance was retrived from the [enterotypes weblink](http://enterotypes.org/ref_samples_abundance_MetaHIT.txt). The missing value was imputed using **KNN** by `impute` package.

```{r}
dat <- read.delim('http://enterotypes.org/ref_samples_abundance_MetaHIT.txt')
dat <- impute::impute.knn(as.matrix(dat), k = 100)
dat <- as.data.frame(dat$data+0.001) 
setDT(dat, keep.rownames = TRUE)
dat
```

Constructe the `bSet` class and then estimate the the proper cluster number using the `estimate_k` function. The `estimate_k` function take advantage of  `Jensen-Shannon divergence` to cluster the samples and the number of clusters was optimizated by Calinski-Harabasz (CH) Index and Silhouette Coefficient. 

The `estimate_k` returns `verCHI` class, a `S3` class containing a optimal cluster results, optimal number cluster, a minmum CHI, a minmum Silhouette value, and Jensen-Shannon divergence matrix. 

```{r}
dat <- bSet(b =  dat)
res <- estimate_k(dat)
res
```

The proper number of cluster is 4. 

Next, the `enterotyping` function was used to identify the enterotype for each cluster and it returns a 3-length list. This list contains two enterotypes matrices and a unidentified samples vector. Cluster 2, 3, and 4 was enterotype Bacteroides, Prevotella, and Ruminococcus, resepectively. 

```{r}
ret=enterotyping(dat, res$verOptCluster) 
ret
```


Furthermore, this result was validated by enterotypes results given by the [enterotype website](http://enterotypes.org).

```{r}
enterotypes <- read.table(system.file('extdata', 'enterotype.txt', package = 'mbOmic'))
enterotypes <- enterotypes[samples(dat),]
table(res$verOptCluster, enterotypes$ET)
```


## SessionInfo

```{r}
devtools::session_info()
```
